Plant Metagenomic Barcoding of Pollen Loads Offers Insights on the Foraging

### 2022-12-04 script/manuscript_MEE.Rmd
more stat values

Sequences were processed using Trimmomatic, which removed sequence adapters, clipped the first 3 bp, discarding reads less than 36 bp, and removing reads if their average PHRED score dropped beneath 20 over a window of 5 bp (@bolger2014trimmomatic). Contigs were generated using HybPiper using target files created by M353 (@johnson2016hybpiper, @mclay2021new).

@@

Using the fuchsin jelly preparation and light microscopic analyses of grains and scoring of 10 character states resulted in the establishment of 28 morphotypes which grains could be reliably classified into. *APPENDIX 7*. 60 Samples were counted and based on rarefaction **had over % of expected morphotypes found** (morphotype richness, mean = 4.5, median = 4, min = 1, max = 9), all samples had expected morphotype diversity reach the asymptote *APPENDIX 8*. The number of counted pollen grains in each sample range from (mean = 2788.685, median = 1453, max = 16293).

@@

XX corbiculae loads had DNA extracted and underwent hyb-seq, a total of XX corbiculae samples were sequenced, XX reads were recovered from sequencing. After trimming XX seqs remained. XX reads were matched using Kraken, of the reads classified by Kraken XX reads were matched using Bracken, of the reads classified by Kraken XX reads were matched using BLAST. The number of reads per sequence varied widely (min = , $\bar{x}$, max = ), the number of loci recovered per sample ranged from (min = , max = ), and the number of reads from loci also ranged widely () **APPENDIX X**.

@@

@@

Only 58 of these 388 species (n = 34.56897, Mdn = 31) were able to be compared to ground truth data from long term (1970 ? - 2016 ?) datasets.

@@

@@

While at the time of writing this there are limited A353 sequence data, the Plant and Fungal Trees of Life (PAFTOL) project, which is sequencing at least a species of each genera in the plant Kingdom will produce sequence data from over 14,000 species. Given the extant publicly available genomic data, we conservatively estimate that upon completion of PAFTOL there will be no fewer than 15,500 species (4.4% of all ca. 350,000 plant species) for which sequence data of a majority of these loci exist (@govaerts2021world). Accordingly, projects in the near future may increase the number of metagenomics samples while decreasing the need to create their own plant sequence reference libraries. 
  As a result of PAFTOL the first ever comprehensive phylogenetic hypotheses of all plant genera will be presented. In tandem with an increased number of digitised and geo-referenced herbarium specimens, and monitoring programs in natural areas, we believe that geo-informatics, and phylogenetic inference will increase the ability of researchers applying this technique to identifying sequence reads. 
  While our approach emphasises the use of this metagenomic technique for the purpose of identifying pollen, I argue the template and resources we provide here make this approach a suitable candidate for many plant metagenomic tasks. While we did not have the resources to explore the possibility of characterising infraspecific characteristics, preliminary results from others (Wenzell et al. in prep., Loke et al. in prep) indicate a possibility for these probes to also collect data at the level of populations and individuals. **

@@

```{r APPENDIX NUMBER LOCI WITH READS USED IN MATCHING, echo = F, fig.height=2.7, fig.width=9}
knitr::include_graphics('../graphics/plots/Percent_loci_matched_by_sample.png')
```

@@

>>>>

Sequences were processed using Trimmomatic, which removed sequence adapters, clipped the first 3 bp, discarding reads less than 36 bp, and removing reads if their average PHRED score dropped beneath 20 over a window of 5 bp (@bolger2014trimmomatic, @tange_2022_6377950). Contigs were generated using HybPiper using target files created by M353 (@johnson2016hybpiper, @mclay2021new).

@@

Using the fuchsin jelly preparation and light microscopic analyses of grains and scoring of 10 character states resulted in the establishment of 28 morphotypes which grains could be classified into. *APPENDIX 7*. 60 samples were counted and based on rarefaction **had over % of expected morphotypes found** (morphotype richness, mean = 4.5, median = 4, min = 1, max = 9), all samples had expected morphotype diversity reach the asymptote *APPENDIX 8*. The number of counted pollen grains in each sample range from ($\bar{x}$ = 2788.685, median = 1453, max = 16,293).

@@

```{r eval = F}
toseq <- read.csv('../data/Benkendorf_combos_fixed.csv') |> 
  filter(!str_detect(extract_id, '[aA-zZ]')) %>% nrow()

krak <- read.csv('../data/kraken_proc_reads.csv')
brack <- read.csv('../data/bracken_proc_reads.csv')
blast <- read.csv('../data/blast_proc_reads.csv')
```

XX corbiculae loads had DNA extracted and underwent hyb-seq, a total of `r toString(toseq)` corbiculae samples were sequenced, XX reads were recovered from sequencing. After trimming XX sequences remained. `r toString(sum(krak[krak$Classified == 'C',]$n))` reads were matched using Kraken, of the reads classified by Kraken `r toString(sum(brack$reads))` reads were matched using Bracken, of the reads classified by Kraken `r toString(sum(blast$Total_seqs))` reads were matched using BLAST. The number of reads per sequence varied widely (min = , $\bar{x}$, max = ), the number of loci recovered per sample ranged from (min = , max = ), and the number of reads from loci also ranged widely () **APPENDIX X**.

@@

```{r, eval = F}
rm(toseq)
```

@@

Only 58 of these 388 species (n = 34.56897, Mdn = 31) were able to be compared to ground truth data from long term (1974–2012) datasets.

@@

A total of 66 corbiculae loads were collected from Bees, 64 of them from Queens.

@@

While at the time of writing this there are limited A353 sequence data, the Plant and Fungal Trees of Life (PAFTOL) project, which is sequencing at least a species of each genera in the plant Kingdom will produce sequence data from over 14,000 species. Given the extant publicly available genomic data, we conservatively estimate that upon completion of PAFTOL there will be no fewer than 15,500 species (4.4% of all ca. 350,000 plant species) for which sequence data of a majority of these loci exist (@govaerts2021world). Accordingly, projects in the near future may increase the number of metagenomics samples while decreasing the need to create their own plant sequence reference libraries. 
  As a result of PAFTOL the first ever comprehensive phylogenetic hypotheses of all plant genera will be presented. In tandem with an increased number of digitised and geo-referenced herbarium specimens, and monitoring programs in natural areas, we believe that geo-informatics, and phylogenetic inference will increase the ability of researchers applying this technique to identifying sequence reads. 
  While our approach emphasises the use of this metagenomic technique for the purpose of identifying pollen, I argue the template and resources we provide here make this approach a suitable candidate for many plant metagenomic tasks. While we did not have the resources to explore the possibility of characterising infraspecific characteristics, preliminary results from others (@wenzell2021incomplete, Loke et al. in prep) indicate a possibility for these probes to also collect data at the level of populations and individuals. **

@@

```{r APPENDIX NUMBER LOCI WITH READS USED IN MATCHING, echo = F}
knitr::include_graphics('../graphics/plots/Percent_loci_matched_by_sample.png')
```

@@

Layer                   Description                            Source
----    ----------------------------------------------  ---------------------
1.             Mean annual cloudiness - MODIS            Wilson et al. 2016
2.           Cloudiness seasonality 1 - MODIS            Wilson et al. 2016 
3.           Cloudiness seasonality 2 - MODIS            Wilson et al. 2016 
4.           Cloudiness seasonality 3 - MODIS            Wilson et al. 2016
5.          Beginning of the frost-free period               Wang et al.
6.             Climatic moisture deficit                     Wang et al.
7.             Degree-days above 5C from                     Wang et al.
8.             Mean annual precipitation                     Wang et al.
9.          Mean annual precipitation as snow                Wang et al.
10.             Temperature seasonality                      Wang et al.
11.       2015 Percent Grass/Herbaceous cover - MODIS         (MOD44B)
12.      2015 Percent Tree cover from Landsat 7/8              (GLCF)
13.       Soil probability of bedrock (R Horizon)             SoilGrids
14.         Soil organic carbon (Tonnes / ha)                 SoilGrids
15.              Surface soil pH in H2O                       SoilGrids
16.             Surface soil percent sand                     SoilGrids
17.                   Soil USDA class                         SoilGrids
18.                Topographic elevation                    EarthEnv DEM
19.          Topographic elevation, moving window.          EarthEnv DEM
20.            Topographic percent slope                    EarthEnv DEM 
21.            Topographic wetness index                    EarthEnv DEM 
22.             Topographic aspect from                     EarthEnv DEM
23.      Annual potential solar radiation computed             r.sun
24.       Estimated actual (w/-cloud) solar radiation  r.sun / Wilson et al. 2016
25.       Log-transformed distance to surface water   Global Surface Water Explorer
26.             Percent surface water                 Global Surface Water Explorer 
----    ----------------------------------------------  ---------------------
